### 2020-04-07 submission/manuscript.Rmd
Incorporate edits from AKH, KLS, BT, and PDS

\vspace{10mm}

@@

Ada K. Hagan${^2}$, Nick Lesniak, *Author names here, order TBD*

Patrick D. Schloss${^1}$${^\dagger}$

@@

We continue to experiment with approaches to running Code Club, but have learned that it is critical for the presenter to clearly articulate their goals and facilitate participant engagement. Although any individual Code Club session may be more experimental than another, on the whole they are a critical tool to train bench scientists in reproducible data analysis practices. We have provided some examples of successful code club topics in Table 1 and summarized the results of our own experimentation as Code Club presenters and participants into Ten Rules. The first 3 rules apply to all participants at every Code Club, while Rules 4 through 8 are targeted to Code Club presenters and the last two for participants.

@@

It is critical that the presenter and participants respect each other and that a designated individual (e.g. the lab director) enforces a code of conduct (see X for an example). Each member of the Code Club must have the humility to acknowledge that they have more to learn about any given topic. We have found that many problems are avoided when the presenter takes charge of the session with a clear lesson plan, thoughtful group creation, and constructive encouragement. Similarly, participants foster a positive environment by remembering that the task is not a competition, focusing on the presenter's goals, allowing their partner to contribute, asking clarifying questions when appropriate, and avoiding distractions (e.g. email, social media). Learning to program is challenging and too often results in a toxic environment of nitpicking and nonconstructive criticism. All parties in a Code Club are responsible for preventing this toxicity by demonstrating respect for themselves and their colleagues.

@@

Part of the humility required to participate in a Code Club is acknowledging that your training is incomplete and that it is possible for you to learn something new. For participants, assume that the presenter has a plan and follow their presented approach. After the Code Club, try to incorporate that material into new code or refactor old code. By practicing the material in a different context you will learn the material better. For presenters, incorporate suggested changes into your code. Either party may identify concepts that they are unsure of, presenting opportunities for further conversation and learning.

@@

As suggested by Rule 7, a significant challenge to presenting at Code Club is selecting topics and activities that appeal to a critical mass of the participants. This is particularly difficult if participants have a wide range of coding experience, such as follows a turnover in group membership. In these circumstances it can be useful to cover commonly used tools but that might result in disinterest of more experienced participants. We have identified several strategies to overcome the challenges presented by participants at varying skill levels. Instead of letting participants form their own pairs, the presenter can select pairs of participants with either similar or differing skill levels, depending on their goals. Partnerships between those with similar skill levels requires the presenter to design appropriate activites each skill level. We have found that commenting code is a good skill for beginners since it forces them to dissect and understand a code chunk line-by-line. An advantage of partnerships with disparate skill levels is assigning a single task, which provides the presenter with a diverse range of methods that achieve the same result. This approach to pairing also helps to graft new members into the group that have emerging programming skills. Regardless of how partners are selected, consider asking the pairs to identify a navigator and a driver [@Bockeler_2020]. The driver types at the computer while the navigator tells them what to type, thus ensuring participation of both partners. Midway through the activity, the presenter can have the partners switch roles. Intentionally forming pairs can also engineer group interactions by breaking up cliques, avoiding potentially disruptive partnerships, or pairing reliable role models with new group members.

@@

It is not possible to fully participate in a Journal Club discussion about a paper that the participant has not read. In that context, coming to Code Club without having installed a dependency is similar to asking a simple question about the Journal Club paper. Both instances show a lack of respect and preparation. Just as a presenter must follow Rule 6 to provide materials ahead of time, participants must review the code in advance, download the data sets, install the necessary packages, and perhaps read up on the topic. If the Code Club is based on a paper or chapters in a data science book (e.g., [@Wickham_2016; @Schloss_2018; @Noble2009]) the participants should read them before the session and consider how they might incorporate the concepts into their own work.

@@

An essential ingredient of any Code Club is active participation from all parties. Having an open laptop on the table and permission to use it can feel like an invitation to get distracted by other work, emails, and browsing the internet; fight that urge and focus on the presenter's goals. Be respectful and allow your partner to contribute. Speak up for yourself and force your partner to let you contribute. If the material seems too advanced for you, it can be frustrating, and tempting to mentally check out. Fortuntately, programming languages like R and Python are generally expressive, which should allow you to engage with the logic, even if the syntax is too advanced. Frankly, understanding the logic of when to use one modeling approach over another is more important than how to use it. If you understand the "why", the "how" will quickly follow. More experienced participants should aim to communicate feedback and coding suggestions at a level that all participants can understand and engage in. Regardless of skill levels, your partner and the presenter put themselves at risk by revealing what they do or do not know. Encourage them, and show your gratitude for helping you learn something new, by fully participating in each Code Club.

@@

All co-authors have been participants in the Code Clubs run as part of PDS's lab meetings. They have had a critical role in shaping the evolution of the sessions. AKH led a lab discussion of Code Clubs to identify the key concepts presented, which were supplemented by contributions from NL. The order of co-authors was determined by...

@@

| **Title**             | **Description**                                    |
| ----------------- | ---------------------------------------------- |
| base vs. ggplot   | Given input data and a figure, recreate the figure using R's base graphics or ggplot syntax   |
| Snakemake         | Given a bash script that contains an analysis pipeline, convert it to a series of recipes (can also be done with GNU Make) |
| DRYing code       | Given script with repeated code, create functions to remove repetition   |
| mothur and Vegan	| Given a pairwise community dissimilarity matrix, compare communities using the `adonis` function in the Vegan R package  |
| tidy data         | Given a wide-formatted data table, convert it to a long, tidy-formatted data table using tools from R's tidyverse  |
| GitFlow           | Participants file and claim an issue to add their name to a README file in a GitHub-hosted repository and file a pull request to complete the issue |
| Integrating R and Google docs   | Scrape a Google docs workbook and clean the data to identify previous Code Club presenters |

>>>>

\vspace*{5mm}

@@

\vspace{25mm}

Ada K. Hagan${^2}$, Nick Lesniak, *Author names here, order alphabetical*, Patrick D. Schloss${^1}$${^\dagger}$

@@

We continue to experiment with approaches to running Code Club. During the period of the Covid-19 pandemic in 2020 when many research labs were shuttered, PDS started hosting weekly Code Club virtually. Participants could engage the material posted to https://www.riffomonas.org/code_club/ via a Zoom meeting or asynchronously using the recorded sessions available on YouTube. Regardless of our approach to leading a Code Club, we have learned that it is critical for the presenter to clearly articulate their goals and facilitate participant engagement. Although any individual Code Club session may be more experimental than another, on the whole they are a critical tool to train bench scientists in reproducible data analysis practices. We have provided some examples of successful code club topics in Table 1. We have summarized the results of our own experimentation as Code Club presenters and participants into Ten Rules. The first 3 rules apply to all participants at every Code Club, while Rules 4 through 8 are targeted to Code Club presenters and the last two for participants.

@@

It is critical that the presenter and participants respect each other and that a designated individual (e.g. the lab director) enforces a code of conduct (see X for an example). Each member of the Code Club must have the humility to acknowledge that they have more to learn about any given topic. We have found that many problems are avoided when the presenter takes charge of the session with a clear lesson plan, thoughtfully creating groups, and giving encouragement. Similarly, participants foster a positive environment by remembering that the task is not a competition, focusing on the presenter's goals, allowing their partner to contribute, asking clarifying questions when appropriate, and avoiding distractions (e.g. email, social media). Learning to program is challenging and too often results in a toxic environment of nitpicking and nonconstructive criticism. All parties in a Code Club are responsible for preventing this toxicity by demonstrating respect for themselves and their colleagues.

@@

Part of the humility required to participate in a Code Club is acknowledging that your training is incomplete and that it is possible for you to learn something new. For participants, assume that the presenter has a plan and follow their presented approach. After the Code Club, try to incorporate that material into new code or refactor old code. By practicing the material in a different context, you will learn the material better. For presenters, incorporate suggested changes into your code. Either party may identify concepts that they are unsure of, presenting opportunities for further conversation and learning.

@@

As suggested by Rule 7, a significant challenge to presenting at Code Club is selecting topics and activities that appeal to a critical mass of the participants. This is particularly difficult if participants have a wide range of coding experience, such as follows a turnover in group membership. In these circumstances it can be useful to cover commonly used tools but that might result in disinterest of more experienced participants. We have identified several strategies to overcome the challenges presented by participants at varying skill levels. Instead of letting participants form their own pairs, the presenter can select pairs of participants with either similar or differing skill levels, depending on their goals. Partnerships between those with similar skill levels requires the presenter to design appropriate activities each skill level. We have found that commenting code is a good skill for beginners since it forces them to dissect and understand a code chunk line-by-line. An advantage of partnerships with disparate skill levels is assigning a single task, which provides the presenter with a diverse range of methods that achieve the same result. This approach to pairing also helps to graft new members into the group that have emerging programming skills. Regardless of how partners are selected, consider asking the pairs to identify a navigator and a driver [@Bockeler_2020]. The driver types at the computer while the navigator tells them what to type, thus ensuring participation of both partners. Midway through the activity, the presenter can have the partners switch roles. Intentionally forming pairs can also engineer group interactions by breaking up cliques, avoiding potentially disruptive partnerships, or pairing reliable role models with new group members.

@@

It is not possible to fully participate in a Journal Club discussion about a paper that the participant has not read. In that context, coming to Code Club without having installed a dependency is similar to asking a simple question about the Journal Club paper. Both instances show a lack of respect and preparation. Just as a presenter must follow Rule 6 to provide materials ahead of time, participants must review the code in advance, download the data sets, install the necessary packages, and perhaps read up on the topic. If the Code Club is based on a paper or chapters in a data science book (e.g., [@Wickham_2016; @Schloss_2018; @Noble2009]), the participants should read them before the session and consider how they might incorporate the concepts into their own work.

@@

An essential ingredient of any Code Club is active participation from all parties. Having an open laptop on the table and permission to use it can feel like an invitation to get distracted by other work, emails, and browsing the internet; fight that urge and focus on the presenter's goals. Be respectful and allow your partner to contribute. Speak up for yourself and force your partner to let you contribute. If the material seems too advanced for you, it can be frustrating, and tempting to mentally check out. Fortunately, programming languages like R and Python are generally expressive, which should allow you to engage with the logic, even if the syntax is too advanced. Frankly, understanding the logic of *when* to use one modeling approach over another is more important than knowing *how* to use it. If you understand the "why", the "how" will quickly follow. More experienced participants should aim to communicate feedback and coding suggestions at a level that all participants can understand and engage in. Regardless of skill levels, your partner and the presenter put themselves in a vulnerable position by revealing what they do or do not know. Encourage them, and show your gratitude for helping you learn something new, by fully participating in each Code Club.

@@

All co-authors have been participants in the Code Clubs run as part of PDS's lab meetings. They have had a critical role in shaping the evolution of the sessions. AKH led a lab discussion of Code Clubs to identify the key concepts presented, which were supplemented by contributions from NL. The order of other co-authors is listed alphabetically.

@@

| **Title**             | **Description**                                    |
| ----------------- | ---------------------------------------------- |
| base vs. ggplot   | Given input data and a figure, recreate the figure using R's base graphics or ggplot syntax   |
| Snakemake         | Given a bash script that contains an analysis pipeline, convert it to a Snakemake workflow (can also be done with GNU Make) |
| DRYing code       | Given script with repeated code, create functions to remove repetition   |
| mothur and Vegan	| Given a pairwise community dissimilarity matrix, compare communities using the `adonis` function in the Vegan R package  |
| tidy data         | Given a wide-formatted data table, convert it to a long, tidy-formatted data table using tools from R's tidyverse  |
| GitFlow           | Participants file and claim an issue to add their name to a README file in a GitHub-hosted repository and file a pull request to complete the issue |
| R with Google docs   | Scrape a Google docs workbook and clean the data to identify previous Code Club presenters |
| Develop an R Package  | Convert a lab member's collection of scripts into package over a series of sessions  |
